%%%%
%% Copyright 2022 Pierre S. Caboche
%% All rights reserved
%%%%


\newpage

\section{Paragraphs, footnotes}

In this section, we will learn how to make paragraphs, sub-paragraphs, etc. \\

The best way to explain how it works is to show the \LaTeX\ code, and show you how it renders. \\



This is the \LaTeX\ code:

\lstinputlisting[language=tex]{content/sample-paragraphs.tex}

\bigskip



\dots and this is the output. Please note that the code explains how paragraphs work. I needed some sample text, but instead of using a generic \LoremIpsum, I figured that this section could be self-describing\dots

\bigskip

\begin{mdframed}
	\input{content/sample-paragraphs.tex}
\end{mdframed}


\bigskip
\bigskip


\subsection{Footnotes} \label{footnote}

In the previous code sample, we saw an example of a \texttt{footnote}. \\

In \LaTeX, it's very easy to add a footnote with the \lstinline[]|\footnote| command:
\begin{lstlisting}[language=tex]
This is some text\footnote{and this is the footnote}.
\end{lstlisting}
\medskip

Footnotes will appear at the foot of the page (see: example\footnote{this is an example of a footnote. As you can see, it is able to hold a lot more information than a margin note
}), 
or at the bottom of components like a \texttt{mdframed} box. \\

As mentioned earlier, footnotes can be useful when you want to add some information but it would be too long to put in parentheses without disturbing the flow of ideas.\\

\bigskip

So the following can be used to add additional information:
\begin{itemize}
	\item parentheses: used for short explanations, closely related to the subject
	\item footnotes: used for additional information that is either too long to be in parentheses, or not closely related to the subject
	\item margin notes (section \ref{margin-notes}): to bring the reader's attention on a particular subject
\end{itemize}

\newpage


\subsection{Margin notes and paragraphs} \label{margin-notes}

\LaTeX\ also allows to put notes in the margins, thanks to \emph{margin paragraph} and \emph{margin notes}.
\\

\idxCmd{marginpar}
\input{content/sample-marginpar.tex}

\medskip

To add a \emph{margin note}, you will need \quotePkg{marginnote} package, 
which provides the \quotePkg{marginnote} command:
\lstinputlisting[language=tex]{include/featured/pkg-marginnote.tex}
\medskip

\idxCmd{marginnote}
\input{content/sample-marginnote.tex}

\bigskip

Notice how in both cases, we put the \quotePkg{marginpar} or \quotePkg{marginnote} command immediately after the first word in the paragraph. This is to make sure the margin paragraph/note will be aligned vertically with the first word of the paragraph. \\

Also keep in mind that in a two-sided document (section~\ref{two-sided}), a margin paragraph will be on the \emph{outer} margin (i.e. left margin for even-numbered pages, right margin for odd-numbered pages). If the document is one-sided, margin paragraphs will appear in the right margin. \\


A \marginpar{marginpar vs. \\ footnote}
margin paragraph (or note) stands out from the rest of the text. 
They can be used to bring the reader's attention on a particular paragraph. However, \emph{it is better to keep them short}. \\

In contrast, a footnote (\emph{section \ref{footnote}}) can be much longer, and is used to provide additional information without breaking the flow of the discussion.
\newpage



\subsection{Putting text in a frame} \label{mdframed}

Previously, we've seen how margin notes and margin paragraphs could be used to bring attention to certain parts of your document. \\

Now we will learn how to make some ideas stand out, by putting them in a big box, thanks to the \quotePkg{mdframed} package:
\lstinputlisting[language=tex]{include/featured/pkg-mdframed.tex}

\bigskip

The \quotePkg{mdframed} package defines, amongst other things, the \quoteEnv{mdframed} environment. Here is a very simple example :
\lstinputlisting[language=tex]{include/featured/example-frame-1.tex}

You can notice how an environment is delimited by \texttt{\textbackslash begin\{\emph{envname}\}} and \\
\texttt{\textbackslash end\{\emph{envname}\}}. \\

And here is the result of the code above:
\input{include/featured/example-frame-1.tex}


\bigskip



This is a very simple frame, but we can modify its appearance with the \quoteCmd{mdfsetup} command.

One thing to know about the \quoteCmd{mdfsetup} command, is that is modifies the properties of \emph{every} \quoteEnv{mdframed} that comes after it. This is a problem if you want to modify the attributes of only \emph{one} frame. \\


To get around this issue, we will create an \emph{environment}. \\
We will name it ``\texttt{vegas}". Why? because \emph{``what happens in \emph{\texttt{vegas}} stays in \emph{\texttt{vegas}}"} (and what happens in an environment stays in that environment"). \\

So here is our ``\texttt{vegas}" environment:
\lstinputlisting[language=tex]{include/featured/example-frame-vegas-env.tex}
\input{include/featured/example-frame-vegas-env.tex}

For the purpose of this example, we declared this environment with one optional parameter: the title of the frame, which defaults to ``!!! Welcome to Vegas !!!". \\

Now it's time to use this environment:
\lstinputlisting[language=tex]{include/featured/example-frame-vegas.tex}

And here is what it looks like: \\
\input{include/featured/example-frame-vegas.tex}

\bigskip

Wow, so many colours! \\


\bigskip

Now, let's create another frame:
\lstinputlisting[language=tex]{include/featured/example-frame-2.tex}

And here is the result:
\input{include/featured/example-frame-2.tex}


As you can see, our very flamboyant ``\texttt{vegas}" environment did not have any influence over to the new frame. The \quoteCmd{mdfsetup} command only affected the environment where it was used (i.e. ``\texttt{vegas}"), and had no effect outside of it. \\

\bigskip


Again, we can define a new environment, call \quoteCmd{mdfsetup} inside of it, and the effect of \quoteCmd{mdfsetup} will not be seen outside of the environment:
\lstinputlisting[language=tex]{include/featured/example-frame-blue-env.tex}
\input{include/featured/example-frame-blue-env.tex}


\lstinputlisting[language=tex]{include/featured/example-frame-blue.tex}


Result:
\input{include/featured/example-frame-blue.tex}


\bigskip

In our two previous environments (``\texttt{vegas}" and ``\texttt{blueframe}"), we passed the frame title as a parameter (whether optional or mandatory). However, if we don't need to add a title, there is a simpler way to define frame environment, thanks to the \quoteCmd{newmdenv} command. \\


Below is the definition for ``\texttt{note}", an environment which I used to add some important notes throughout this document, in the form of a frame with a slightly yellow background:
\lstinputlisting[language=tex]{include/featured/note-env.tex}

Here is how to call it:
\lstinputlisting[language=tex]{include/featured/example-frame-note.tex}
Here is the result:
\input{include/featured/example-frame-note.tex}


\bigskip

As you can see, all the options that we would normally pass to the \quoteCmd{mdfsetup} command are passed in the optional parameter (in square brackets \texttt{[]}) of \quoteCmd{newmdenv}. \\
The name of the environment is passed as the mandatory argument (in curly brackets \texttt{\{\}}) of \quoteCmd{newmdenv}.

\bigskip

Another way to modify the appearance of a \quoteEnv{mdframed} is to define a new style with the \quoteCmd{mdfdefinestyle} command, then use that style by specifying the \texttt{style} attribute in the \quoteEnv{mdframed} options:

\lstinputlisting[language=tex]{include/featured/example-frame-pinky-env.tex}
And the result:
\input{include/featured/example-frame-pinky-env.tex}

\bigskip


This concludes our presentation of \quotePkg{mdframed}. \\

If you want to know more about this package (with many other examples), please refer to its official documentation: 
\begin{center}
\url{https://www.ctan.org/pkg/mdframed}
\end{center}


\newpage

\subsection{Other types of boxes}

There exists other kinds of boxes, to draw boxes around text:
\begin{itemize}
	\setlength{\itemsep}{-0.5em}
	
	\item \quoteCmd{parbox}
	\item \quoteCmd{mbox}
	\item \quoteCmd{makebox}
	\item \quoteCmd{fbox}
	\item \quoteCmd{framebox}
	\item \quotePkg{fancybox} (package)
\end{itemize}

\bigskip

For more information, please see:
\begin{center}
	\url{https://en.m.wikibooks.org/wiki/LaTeX/Boxes}
\end{center}

\bigskip



\section{Comments} \label{comments}


In \LaTeX, a comment starts with: \% \\


Everything after the ``\%" till the end of the line will be ignored. If you want to actually print a ``\%" in \LaTeX, you'll need to escape it with a backslash: \textbackslash \% \\

Unlike in other computer languages, there are no multiline comments in \LaTeX, only single-line comments.
\\


Fortunately, an editor like TeXStudio can make the process of adding/removing comments a lot easier, thanks to the \texttt{Ctrl+T} shortcut (for \emph{``Toggle Comment"}).

\begin{mdframed}
\medskip
	To toggle comments in TeXStudio:
	\begin{itemize}
		\item Select the lines you want to add comments to (or remove comments from)
		\item Press \texttt{Ctrl+T}
		
		This will either add or remove the comments from those lines
	\end{itemize}
\medskip
\end{mdframed}



Comments are extremely useful in \LaTeX! In fact\dots 

\begin{note}
When you think about it, comments are one of the most powerful features in \LaTeX. \\

Indeed, comments allow you to keep multiple drafts of the same paragraphs, which you can then enable/disable with a single \texttt{Ctrl+T} in \TeXstudio. This way, you never have to throw away any idea! (just comment them out, so they don't show up in the document).
\end{note}



Comments have a number of applications:

\begin{description}
	\item[\emph{Describing}] \mbox{}\\
	The most obvious use of a comment is to describe what some \LaTeX\ code is doing (e.g. describe the packages that you are using and what you are using them for)
	
	\item[\emph{Taking notes}] \mbox{}\\
	For example, if you already have your document's plan laid out (i.e. you've already defined the structure in terms of: parts, sections, sub-sections\ldots), you can leave comments to list the main ideas to put in those sections later (as some kind of \emph{TODO} list)
	
	Furthermore, if several people are working on a document you can leave notes to other contributors in the form of comments (and if you work alone, you can always leave notes to yourself)
	
	\item[\emph{Drafting ideas}] \mbox{}\\
	This is probably the most important use of comments (this represents about 99\% of the cases when I use comments in \LaTeX)
	
	Sometimes you're writing things but are not completely satisfied with the wording. In \LaTeX, it's very easy to put a sentence in comments, rewrite it, and repeat the process until you're satisfied (while not having to discard what you wrote, and not having those earlier versions appear in your document).
\end{description}


In many of those cases, you don't want your comments to be removed by accident (i.e. any case other than \emph{Drafting ideas}). To avoid that, simple use more than one ` \lstinline|%| ' for your comment. \\

This way, one \texttt{Ctrl+T} will not be enough to remove the comment\ldots
\begin{lstlisting}[language=tex]
%%% Packages and configuration
\end{lstlisting}


\bigskip


\section{Collaboration, version management}

The main focus of \LaTeX\ is document generation. As such, it does not provide dedicated collaboration and version management tools. \\

However, \LaTeX\ is a markup language. This means that, collaboration and management of \LaTeX\ files can be handled with the same tools as any other source files; so you can use your favourite version management tools with \LaTeX\ (e.g. GIT, SVN\dots) \\

In section ``\longref{comments}", we've seen how \LaTeX\ comments can be used for note-taking, drafting entire paragraphs, etc. A version-control management system can help you keep a history of such items.


\newpage

\part{Managing breaks}

In part \ref{your-first-document}, we talked about how to how to write paragraphs, how to organise your documents in sections, etc. \\

Sometimes, however, a paragraph or section break does not land in a place that is convenient for us, as it disrupts the flow of the article, For example, a section may start at the bottom of a page, a figure may end up on a different page from the text that references it (I always find this confusing to read when that happens), etc. 

To avoid that, you might want to make some manual adjustments to the document layout. This is the subject of this part\dots

\section{Non-breaking spaces}

In \LaTeX, you can create a non-breaking space by using a tilda: $\sim$ \\

Non-breaking spaces are used to keep certain terms together, without automatic line breaks being inserted in the middle. \\

This is useful, for example, to keep numbers and their units together:

\begin{lstlisting}[language=tex]
	The maximum theoretical speed of SATA-3 is 6~Gb/s, or 750~MB/s.
\end{lstlisting}

\medskip

Whenever possible, \LaTeX\ will keep those terms together. \\

In the previous example, the ``6" should not be separated from its ``Gb/s" unit, and the ``750" should not be separated from the ``MB/s".

\bigskip


\section{Phantom text} \label{phantom}

Since we are on the subject of adding breaks and spaces, here are the commands to add some empty space based on the width/height of a given text:

\begin{description}
	\item{\quoteCmd{phantom}\texttt{\{text\}}} \mbox{} \\ 
		Creates some space, of the same width and height as the \texttt{text} passed as parameter
		(a combination of \quoteCmd{hphantom} and \quoteCmd{vphantom})
		
	\item{\quoteCmd{hphantom}\texttt{\{text\}}} \mbox{} \\ 
		Creates some horizontal space, of the same width as the \texttt{text} passed as parameter
	\item{\quoteCmd{vphantom}\texttt{\{text\}}} \mbox{} \\ 
		Creates some vertical space, of the same height as the \texttt{text} passed as parameter
\end{description}

\newpage

Below is an example of how to put an \quoteCmd{underline} under some \quoteCmd{phantom} text:
\begin{lstlisting}[language=tex]
I really think that \underline{\phantom{\LaTeX}} is great!
\end{lstlisting}

And here is the result:
\begin{mdframed}
I really think that \underline{\phantom{\LaTeX}} is great!
\end{mdframed}

\bigskip

\section{Skips and breaks}


\subsection{Adding vertical spaces}

Sometimes you might need to add some extra vertical space between paragraphs (or other elements). These are called ``skip". \\

Skips come in different sizes: \\

\begin{tabular}{l l}
	\quoteCmd{smallskip} & a small skip \\
	\quoteCmd{medskip}   & a medium skip \\
	\quoteCmd{bigskip}   & a big skip \\
\end{tabular}

\bigskip


\subsection{Page breaks}

You can start a new page with the \quoteCmd{newpage} command. \\

You can also start a new page with \quoteCmd{pagebreak}, in which case the content of the old page will be spread out vertically. \\

Having several \quoteCmd{newpage} (or \quoteCmd{pagebreak}) will not create empty pages. To insert empty page(s), see section \ref{blank-page}\dots

\bigskip


\subsection{Inserting a blank page} \label{blank-page}


While writing about \quoteCmd{newpage} and \quoteCmd{pagebreak}, I wondered \emph{``how to insert a blank page in a \LaTeX\ document?"} \\

After a minute of searching, I found the following solution, which I then wrapped in a custom macro with an optional parameter (section \ref{macro-optional-parameter}):


\lstinputlisting[language=tex]{include/featured/cmd-insertBlankPage.tex}


The macro will insert a blank page, with a custom message (passed as an optional parameter) written in the center of the page. If the message is not specified, it will default to ``This page intentionally left blank." \\

This macro uses the \quoteCmd{thispagestyle} command, which sets the style for the current page only. By setting it to \texttt{empty}, we remove the header and footer temporarily. Therefore only the message will appear on the center of the page. \\

Here are some examples of how to call our custom macro:
\input{include/featured/cmd-insertBlankPage-example.tex}

\bigskip

We will be calling the following:
\begin{lstlisting}
\insertBlankPage
\end{lstlisting}

See next page for the result\dots
\insertBlankPage



